```{r 02-setup, warning=F,message=F}
library(tidyverse)
library(tidyr)
library(ggthemes)
library(lubridate)

# Now that we have learned how to munge (manipulate) data
# and plot it, we will work on using these skills in new ways
# 
# knitr::opts_knit$set(root.dir='..')
```


# Hayman Fire Recovery

```{r 02-dataread, warning=F,message=F}
####-----Reading in Data and Stacking it ----- ####
#Reading in files
files <- list.files("./02-data", full.names=T)
# If run from line, lowest directory: need ../ | If run from console, base directory at .Rproj

paste(files)

#Read in individual data files
ndmi <- read_csv(files[1]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndmi')

ndsi <- read_csv(files[2]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndsi')

ndvi <- read_csv(files[3])%>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndvi')

# Stack as a tidy dataset
full_long <- rbind(ndvi,ndmi,ndsi) %>%
  gather(key='site',value='value',-DateTime,-data) %>%
  filter(!is.na(value))


```


## Question 1) 

What is the correlation between NDVI and NDMI? - here I want you to
convert the full_long dataset in to a wide dataset using the 
function "spread" and then make a plot that shows the correlations a
function of if the site was burned or not (x axis should be ndmi)
You should exclude winter months and focus on summer months

```{r}
wide_long <- spread(full_long, key = "data", value = "value") %>% 
  mutate(month = month(DateTime)) %>% 
  filter(month %in% c(5:9))

ggplot(wide_long, aes(ndmi, ndvi, color = site)) + geom_point()
```


## Question 2 

2) What is the correlation between average NDSI (normalized
 snow index) for January - April and average NDVI for June-August?
In other words, does the previous year's snow cover influence vegetation
 growth for the following summer?

```{r}
# Group by year (early NDVI  /  late NDSI). Compare the different years, not months ##Got help with this question on Monday, did not have the chance to complete it (02-07-2022)

# Filter and prepare data to be aggregated by month, year.

wide_early_ndsi <- spread(full_long, key = "data", value = "value") %>%
  mutate(month = month(DateTime)) %>%
  mutate(year = year(DateTime)) %>% 
  filter(month %in% c(1:4)) %>% 
  select(site, ndsi, month, year) %>% 
  group_by(month, year)

wide_late_ndvi <- spread(full_long, key = "data", value = "value") %>% 
  mutate(month = month(DateTime)) %>% 
  mutate(year = year(DateTime)) %>% 
  filter(month %in% c(6:8)) %>% 
  select(site, ndvi, month, year) %>% 
  group_by(month, year)

# Aggregate data values for ndsi/ndvi by year

ndsi_yearly <- aggregate(ndsi ~ year + site, wide_early_ndsi, FUN = mean)
  
ndvi_yearly <- aggregate(ndvi ~ year + site, wide_late_ndvi, FUN = mean)

# Combine data for plotting

total <- full_join(ndsi_yearly, ndvi_yearly)

# Plot of the data, color by type (ndvi, ndsi)

ggplot() + geom_point(data = filter(total, site == "burned"), aes(year, ndsi, col = "ndsi")) +
  geom_point(data = filter(total, site == "unburned"), aes(year, ndvi, col = "ndvi")) +
  scale_x_continuous(n.breaks = 20) + scale_color_manual(values=c("cyan3", "chartreuse4"))

cor(total$ndsi, total$ndvi)
```

Snowfall has a weak short-term impact on the later seasons vegetation index. We can see in the graph that years with lower snow indexes over a longer period cause a small change in the NDVI in the future years.



## Q3

How is the snow effect from question 2 different between pre- and post-burn
and burned and unburned? 


Significantly lower snow is detected in the initial post-burn area. From the graph, we can see that the index values drop significantly initially, the recover over the course of about 4 years.

## Question 4

What month is the greenest month on average? 

```{r}

# Month with highest NDVI on average, rename months

ndvi_monthly <- aggregate(ndvi ~ month, wide_late_ndvi, FUN = mean) %>% 
  mutate(month = month(month, label = TRUE, abbr = FALSE))

# Find max index
max_ndvi <- which.max(ndvi_monthly[,2])

# Print max index month name
print(paste(ndvi_monthly[max_ndvi,1], "is the greenest month on average."))

```


## Question 5) 

What month is the snowiest on average?

```{r}

# Month with highest NDSI on average, rename months

ndsi_monthly <- aggregate(ndsi ~ month, wide_early_ndsi, FUN = mean) %>% 
  mutate(month = month(month, label = TRUE, abbr = FALSE))

# Find max index
max_ndsi <- which.max(ndsi_monthly[,2])

# Print max index month name
print(paste(ndsi_monthly[max_ndsi,1], "is the snowiest month on average."))

```


